---
title: 'STA304'
author: "Neil Montgomery"
date: "2016-02-22"
output: 
  ioslides_presentation: 
    css: 'styles.css' 
    widescreen: true 
    transition: 0.001
---
\newcommand{\E}[1]{E{\left(#1\right)}}
\newcommand{\flist}[2]{\{#1_1, #1_2, \ldots, #1_#2\}}
\newcommand{\fulist}[3]{\{#1_{{#2}1}, #1_{{#2}2}, \ldots, #1_{{#2}{#3}}\}}
\renewcommand{\bar}[1]{\overline{#1}}
\newcommand{\SE}[1]{\sqrt{\hat{V}(#1)}}



# stratified random sampling

## recap

* Stratified sampling can be useful for:
    + getting better population parameter estimates, if the strata are (more) homogeneous
    + further investigation of strata
    
* SRS done within each strata to get the usual estimates with the usual properties

* Population paramaters are estimated using suitably weighted combinations of stratum estimates. 

## (from last time)

Going backwards: $\hat\tau = N\bar y_{st}$.

True and estimated variance (because of independence):
$$\begin{align*}
V(\bar y_{st}) &= \sum_{i=1}^L W_i^2 V(\bar{y}_i)\\
\hat V(\bar y_{st}) &= \sum_{i=1}^L W_i^2 \hat V(\bar{y}_i)
\end{align*}$$

Confidence interval $\bar y_{st} \pm 2\SE{y_{st}}$, bound on error of estimation, etc.

## example(s)

```{r, echo=FALSE, message=FALSE}
library(rio)
library(dplyr)
tx <- import("tx.csv")
N <- nrow(tx)
n <- 600
```

We will look at the transformer "data" from the term test, by which I mean the actually simulated population's worth of data that was used for the test, contained in the file `tx.csv`. The story here has to be a little different. On the test knew only the locations of the transformer. For this example we'll pretend we know, say, the location and the `Size` (50KVA, 75KVA, or 100KVA) but not the Manufacturer or the Age.  

## plan: compare SRS versus stratified by `Size`. 

We will try to estimate the average age of the population of transformers. 

We'll do this in two ways. One is using a SRS of size $n = `r n`$. For the other we'll stratify by transformer "Size", which is one of 50KVA, 75KVA, or 100KVA. A summary of the population by Size is as follows, including weights to be used in the stratified formulae later on. Note that here "Size" is a property of a transormer and not anything to do with how many of them there are.

```{r, echo=FALSE}
library(knitr)
tx %>% 
  group_by(Size) %>% 
  summarise(N=n(), W = n()/nrow(tx)) -> N_by_size
kable(N_by_size)
```

## the simple random sample

We select a simple random sample of `r n` transformers and get the following sample mean, standard deviation, and "bound on the error of estimation"

```{r, echo=FALSE}
set.seed(1)
tx_srs <- sample_n(tx, 600)

tx_srs %>% 
  summarize(mean = mean(Age), sd = sd(Age), 
            B = 2*sqrt(var(Age)/n*(N-n)/N)) -> tx_srs_est

kable(tx_srs_est)
```

## stratify by `Size`

To make a fair comparison we'll keep an overall sample size of `r n`. We will choose to allocate the sample proportionally to the size of the strata (more on this later). Here is a summary of the results by stratum:
```{r, echo=FALSE}
set.seed(2)
tx %>% 
  group_by(Size) %>% 
  sample_frac(n/N) -> tx_bySize

tx_bySize %>% 
  group_by(Size) %>% 
  summarise(n = n(), means = mean(Age), variances = var(Age), sds = sd(Age)) -> strat_summ
kable(strat_summ)
```

## stratified estimates

The stratified estimator for the population mean is:

$$ \bar y_{st} = \sum_{i=1}^L W_i\bar{y}_i $$

with estimated variance:

$$ \hat V(\bar y_{st}) = \sum_{i=1}^L W_i^2 \hat V(\bar{y}_i)$$

where $\hat V(\bar{y}_i) = \frac{s^2}{n_i}\frac{N_i-n_i}{N_i}$.

## stratified estimates

Plug in numbers and weights from previous slides ago to get:

$$\bar y_{st} = `r sum(strat_summ$means * N_by_size$W)`$$
$$\hat V(\bar y_{st}) = `r sum(strat_summ$variances/strat_summ$n * (N_by_size$N - strat_summ$n)/N_by_size$N * N_by_size$W^2)`$$

The "usual bound on the error of estimation" is $\sqrt{\hat V(\bar y_{st})} = `r 2*sqrt(sum(strat_summ$variances/strat_summ$n * (N_by_size$N - strat_summ$n)/N_by_size$N * N_by_size$W^2))`$.

